A Bayesian approach to uncertainty in word embedding bias estimation

### 2023-12-13 paper/CLsubmission/bayesBiasCL.tex
applied final reviewers' suggestions to bayesBiasCL.tex

@@

\item \citet{may-etal-2019-measuring} develop a generalization of \textsf{WEAT} meant to apply to
  sets of sentences, \textsf{SEAT}, which basically applies
  \textsf{WEAT} to vector representations of sentences. The authors,
  however, still pre-average and play the game of finding a
  single-number metric, so our remarks apply.

@@

\item \citet{Ethayarajh2020Bernstein} argues that a bias estimate should not be expressed as a single number without taking into account that the estimate is
made using a sample of data and therefore has intrinsic uncertainty.
The author suggest using Bernstein bounds to gauge the uncertainty in
terms of confidence intervals. We do not discuss this approach
extensively, as we think that confidence intervals are quite
problematic for several reasons, among others the confusing
interpretability. We do not think that Bernstein bounds provide the
best solution to the problem. Applying this method to a popular
WinoBias dataset leads to the conclusion that more than 11903 data points of protected words 
are needed to claim a 95\% confidence interval for a bias estimate.
This amount vastly exceeds the existing word lists for bias
estimation. We propose a more realistic Bayesian method. Our
conclusion is still that the word lists are sometimes too small, but
at least they allow for gauging uncertainty as we go on to improve our
methodology and extend the lists gradually.

@@

\item \citet{Spliethover2021BiasSilhouette} propose Bias Silhouette Analysis (BSA), a method for assessing the quality of metrics that measure bias in word embedding models based on word lists. The core idea here is to quantify how much the bias values of a metric vary depending on what words from the lists are actually observed, where the computations result in  values for each model obtained  using word list subsets of increasing length. This allows for an inspection of bias metric convergence and sensitivity to word list choice, with a biased (GloVe) and an explicitly debiased model whose lower bias has been confirmed empirically (NBatch).  They examine the Embedding Coherence Test (ECT), the Relative Negative Sentiment Bias (RNSB), and the Word Embedding Association Test (WEAT), concluding  that none of these metrics can reliably discriminate between biased and non-biased models in all cases. This is not unexpected. An interesting question is what would happen if a similar test was applied to our method. However, our point is that the existing word lists are too short to provide reliable estimate of bias.

>>>>

Recent research, however, has criticized the existing methods for assessing bias in word embedding models. \citet{zhang2020robustness} highlight the limitations of word pairs in detecting bias, arguing that analogies may not necessarily capture societal bias, but instead reflect co-occurrence frequency. \citet{Ethayarajh2019understanding} draw attention to potential problems with the Word Embedding Association Test (WEAT) in degenerate cases, emphasizing its sensitivity to word frequency and list selection. Similarly, \citet{schroder2021evaluating} critique existing bias metrics such as \textsf{MAC} and \textsf{WEAT} in light of their failure to satisfy plausible general formal principles.  \citet{Goldfarb2021BiasNotCorrelate} find only limited correlation between intrinsic bias as measured by WEAT and extrinsic metrics such as equality of opportunity and predictive parity. \citet{Du2021Assessing} discover inconsistencies among existing bias measures, indicating sensitivity to embedding algorithms and corpora. Finally, \citet{Spliethover2021BiasSilhouette} introduces Bias Silhouette Analysis (BSA) as a method for assessing the quality of metrics that measure bias in word embedding models based on word lists, and concludes that none of the metrics they evaluated can reliably discriminate between biased and non-biased models in all cases.  \citet{Ethayarajh2020Bernstein} argue against expressing bias as a single number without considering the inherent uncertainty in sample-based estimates. \citet{Lum2022Debiasing} proposes a double-corrected variance estimator for unbiased estimates and uncertainty quantification of group-wise performance metrics. Both proposals enrich the existing measures with classical statistical uncertainty quantification. Our proposal is to get rid of the single-number measures and use Bayesian methods.

@@

\item \citet{may-etal-2019-measuring} develop a generalization of \textsf{WEAT} meant to apply to
  sets of sentences, \textsf{SEAT}, which basically applies
  \textsf{WEAT} to vector representations of sentences. The authors,
  however, still pre-average and focus on a
  single-number metric, so our remarks apply.

@@

\item \citet{Ethayarajh2020Bernstein} argues that a bias estimate should not be expressed as a single number without taking into account that the estimate is
made using a sample of data and therefore has intrinsic uncertainty.
The author suggests using Bernstein bounds to gauge the uncertainty in
terms of confidence intervals. We do not discuss this approach
extensively, as we think that confidence intervals are quite
problematic for several reasons, among others the confusing
interpretability. We do not think that Bernstein bounds provide the
best solution to the problem. Applying this method to a popular
WinoBias dataset leads to the conclusion that more than 11903 data points of protected words 
are needed to claim a 95\% confidence interval for a bias estimate.
This amount vastly exceeds the existing word lists for bias
estimation. We propose a more realistic Bayesian method. Our
conclusion is still that the word lists are sometimes too small, but
at least they allow for gauging uncertainty as we go on to improve our
methodology and extend the lists gradually.

@@

\item \citet{Spliethover2021BiasSilhouette} propose Bias Silhouette Analysis (BSA), a method for assessing the quality of metrics that measure bias in word embedding models based on word lists. The core idea here is to quantify how much the bias values of a metric vary depending on what words from the lists are actually observed, where the computations result in  values for each model obtained  using word list subsets of increasing length. This allows for an inspection of bias metric convergence and sensitivity to word list choice, with a biased (GloVe) and an explicitly debiased model whose lower bias has been confirmed empirically (NBatch).  They examine the Embedding Coherence Test (ECT), the Relative Negative Sentiment Bias (RNSB), and the Word Embedding Association Test (WEAT), concluding  that none of these metrics can reliably discriminate between biased and non-biased models in all cases. This is in line with our results. An interesting question is what would happen if a similar test was applied to our method. However, our point is that the existing word lists are too short to provide reliable estimate of bias.
